Reanalysis Step 3: Propensity Score Analysis with Twang

### 2025-08-17 reanalysis-3-propensity_scores.qmd
cobalt
twang

# Additional packages
library(twang)
library(survey)
library(cobalt)
library(ggplot2)
library(gridExtra)
library(knitr)

@@

```{r ps-distribution}
#| fig-height: 8
#| fig-width: 12

@@

  # Create data for plotting
  plot_data <- data.frame(
    ps = twang_results[[i]]$ps_scores,
    treatment = factor(imputed_datasets[[i]]$ever_lapse_binary,
                      levels = c(0, 1),
                      labels = c("Control", "Treated"))
  )

  # Histogram of propensity scores by treatment group
  p1 <- ggplot(plot_data, aes(x = ps, fill = treatment)) +
    geom_histogram(alpha = 0.6, position = "identity", bins = 30) +
    labs(title = paste("Twang GBM Propensity Scores - Imputation", i),
         x = "Propensity Score",
         y = "Count") +
    theme_minimal() +
    scale_fill_manual(values = c("blue", "red")) +
    facet_wrap(~ treatment, ncol = 1, scales = "free_y")

  # Density plot overlay
  p2 <- ggplot(plot_data, aes(x = ps, color = treatment, fill = treatment)) +
    geom_density(alpha = 0.3) +
    labs(title = paste("Density of Propensity Scores - Imputation", i),
         x = "Propensity Score",
         y = "Density") +
    theme_minimal() +
    scale_fill_manual(values = c("blue", "red")) +
    scale_color_manual(values = c("blue", "red"))

  # Boxplot comparison
  p3 <- ggplot(plot_data, aes(x = treatment, y = ps, fill = treatment)) +
    geom_boxplot(alpha = 0.6) +
    geom_jitter(width = 0.1, alpha = 0.2, size = 0.5) +
    labs(title = paste("Distribution Comparison - Imputation", i),
         x = "Treatment Group",
         y = "Propensity Score") +
    theme_minimal() +
    scale_fill_manual(values = c("blue", "red"))

  # Arrange plots
  grid.arrange(p1, p2, p3, ncol = 3)
}
```

>>>>

# Additional packages
library(twang)
library(ggplot2)
library(gridExtra)  # For arranging multiple plots
library(knitr)

@@

```{r ps-distribution}
#| fig-height: 6
#| fig-width: 10

@@

  # Get propensity scores from the twang model
  twang_model <- twang_results[[i]]$model
  weights <- twang_results[[i]]$weights
  treated <- imputed_datasets[[i]]$ever_lapse_binary == 1

  # Extract propensity scores - twang stores them in the ps slot
  ps_scores <- tryCatch({
    if (!is.null(twang_model$ps)) {
      # ps can be a matrix (multiple stop methods) or vector (single stop method)
      if (is.matrix(twang_model$ps) || is.data.frame(twang_model$ps)) {
        twang_model$ps[, 1]  # Get first column
      } else {
        twang_model$ps  # Already a vector
      }
    } else {
      # Shouldn't happen but fallback to weights
      weights
    }
  }, error = function(e) {
    weights  # Use weights as fallback
  })

  # Create simple histograms
  par(mfrow = c(1, 2))

  # Control group
  hist(ps_scores[!treated], 
       main = paste("Control - Imputation", i),
       xlab = "Propensity Score",
       col = "lightblue",
       breaks = 20)

  # Treated group  
  hist(ps_scores[treated],
       main = paste("Treated - Imputation", i),
       xlab = "Propensity Score", 
       col = "lightcoral",
       breaks = 20)

  par(mfrow = c(1, 1))
}
```
